---
title: "homework2-506"
format: html
editor: visual
---
This is a link for hw2: https://github.com/lizhixuan202603-sudo/homework2-506


#Setup
```{r}
# Load required packages
library(microbenchmark)   # Problem 1f
library(dplyr)            # Data manipulation
library(ggplot2)          # Plotting

# Set a global seed for reproducibility
set.seed(506)
```

#Problem 1.
#1a. Simulation function
## Function momentum_walk(n, p)
```{r}
# Random Walk with Momentum in 2D

# Simulates a particle moving on a 2D grid. The first step is chosen
# uniformly from four directions; each subsequent step preserves the
# previous direction with probability `p`, and otherwise chooses a new
# direction uniformly from the four directions.
#
# @param n Positive integer. Number of steps (n >= 1).
# @param p Numeric. Probability of preserving the previous direction (0 <= p <= 1).
# @return A numeric vector of length 2, c(x, y), the final coordinates.
# @examples
# momentum_walk(10, 0.6)
momentum_walk <- function(n, p) {
  # Input validation
  if (!is.numeric(n) || length(n) != 1 || is.na(n)) stop("n must be a single numeric value")
  if (!is.numeric(p) || length(p) != 1 || is.na(p)) stop("p must be a single numeric value")
  if (n < 1 || n != as.integer(n)) stop("n must be an integer >= 1")
  if (p < 0 || p > 1) stop("p must be in [0, 1]")

  # Four directions: right, left, up, down
  dirs <- matrix(c(1, 0,
                  -1, 0,
                   0, 1,
                   0,-1), ncol = 2, byrow = TRUE)

  # First step: uniform choice
  d <- sample(1:4, 1)
  x <- dirs[d, 1]; y <- dirs[d, 2]
  prev <- d

  # Subsequent steps
  if (n > 1) {
    for (i in 2:n) {
      if (runif(1) < p) {
        d <- prev                       # keep previous direction
      } else {
        d <- sample(1:4, 1)             # choose a new random direction
      }
      x <- x + dirs[d, 1]
      y <- y + dirs[d, 2]
      prev <- d
    }
  }
  c(x, y)
}

momentum_walk(10, 0.6)
```

##ii. Input sanitization
##The function above already validates n >= 1 (and integer) and 0 <= p <= 1, raising an error otherwise. Test:
```{r}
# Test invalid inputs
tryCatch(momentum_walk(0, 0.5), error = function(e) message(e$message))    # n < 1
tryCatch(momentum_walk(10, 1.5), error = function(e) message(e$message))   # p > 1
tryCatch(momentum_walk(10.5, 0.5), error = function(e) message(e$message)) # non-integer
```

#1b. Two implementations
##Version 1: for loop — momentum_walk1(n, p)
```{r}
# Random Walk with Momentum — for-loop implementation
#
# @param n Number of steps.
# @param p Probability of preserving direction.
# @return Final coordinates c(x, y).
momentum_walk1 <- function(n, p) {
  if (!is.numeric(n) || length(n) != 1 || is.na(n)) stop("n must be a single numeric value")
  if (!is.numeric(p) || length(p) != 1 || is.na(p)) stop("p must be a single numeric value")
  if (n < 1 || n != as.integer(n)) stop("n must be an integer >= 1")
  if (p < 0 || p > 1) stop("p must be in [0, 1]")

  dirs <- matrix(c(1, 0, -1, 0, 0, 1, 0, -1), ncol = 2, byrow = TRUE)

  # First step
  d    <- sample(1:4, 1)
  step <- dirs[d, ]
  x <- step[1]; y <- step[2]

  # Step-by-step loop
  for (i in 2:n) {
    if (runif(1) < p) {
      d <- d
    } else {
      d <- sample(1:4, 1)
    }
    step <- dirs[d, ]
    x <- x + step[1]
    y <- y + step[2]
  }
  c(x, y)
}

#Logic: The for loop advances the particle step-by-step. At each step it first decides whether to keep the previous direction, then updates the coordinates. It is intuitive but loops in R are relatively slow.
```


##Version 2: sapply implementation — momentum_walk2(n, p)
```{r}
# Random Walk with Momentum — sapply implementation

# @param n Number of steps.
# @param p Probability of preserving direction.
# @return Final coordinates c(x, y).
momentum_walk2 <- function(n, p) {
  if (!is.numeric(n) || length(n) != 1 || is.na(n)) stop("n must be a single numeric value")
  if (!is.numeric(p) || length(p) != 1 || is.na(p)) stop("p must be a single numeric value")
  if (n < 1 || n != as.integer(n)) stop("n must be an integer >= 1")
  if (p < 0 || p > 1) stop("p must be in [0, 1]")

  dirs <- matrix(c(1, 0, -1, 0, 0, 1, 0, -1), ncol = 2, byrow = TRUE)

  # Pre-generate: keep[i] = whether step i reuses previous direction;
  #               rand[i] = a pre-drawn random direction for step i.
  keep <- c(TRUE, runif(n - 1) < p)        # step 1 has no "keep" concept; placeholder TRUE
  rand <- sample(1:4, n, replace = TRUE)   # a pre-drawn direction per step

  # Use sapply to determine the actual direction at each step
  idx <- integer(n)
  idx[1] <- rand[1]
  sapply(2:n, function(i) {
    if (keep[i]) {
      idx[i] <<- idx[i - 1]
    } else {
      idx[i] <<- rand[i]
    }
    idx[i]
  })

  # Accumulate coordinates in one vectorized step
  xy <- dirs[idx, , drop = FALSE]
  c(sum(xy[, 1]), sum(xy[, 2]))
}

#Logic: All "keep" decisions and candidate directions are generated vectorially first; then sapply sequentially resolves each step's direction; finally coordinates are summed in one shot with colSums, avoiding an explicit for loop.

```


```{r}
# Demonstration
set.seed(1)
momentum_walk1(10, 0.6)
momentum_walk2(10, 0.6)
momentum_walk1(1000, 0.6)
momentum_walk2(1000, 0.6)
```

#1c.Reproducibility
```{r}
# Demonstrate reproducibility
set.seed(506)
r1 <- momentum_walk1(10, 0.6)
r2 <- momentum_walk1(10, 0.6)   # no reseed → different result
set.seed(506)
r3 <- momentum_walk1(10, 0.6)   # reseed → same as r1

c(r1, r2, r3)   # r1 == r3
```


```{r}
# Same demonstration for n = 1000
set.seed(506)
m1 <- momentum_walk1(1000, 0.6)
set.seed(506)
m2 <- momentum_walk1(1000, 0.6)
identical(m1, m2)   # TRUE

#Explanation: set.seed() sets the internal state of the pseudo-random number generator. With the same seed, all subsequent random calls produce an identical sequence, so results are reproducible. This is a necessary habit for simulation studies.
```

#1d. Does momentum matter?
##i. One simulation for each p
```{r}
set.seed(506)
p_vals <- c(0, 0.5, 0.9)
res_d  <- sapply(p_vals, function(pp) {
  pos <- momentum_walk1(1000, pp)
  sqrt(sum(pos^2))     # distance D
})
names(res_d) <- paste0("p=", p_vals)
print("Example output: ")
res_d
```

##ii. Larger momentum → larger final distance
From the output, the final distance D clearly increases as p increases.

##iii. Reasoning
When p = 0, each step is independent, so the particle performs a classical 2D random walk; the variance of displacement grows linearly with steps, and the final distance is roughly √n. When p approaches 1, the particle keeps moving in nearly the same direction, so displacement grows nearly linearly (~n), and thus the final distance is much larger than a random walk. Higher momentum means more "straight-line-like" motion, so cancellations are less likely.

#1e. Monte Carlo simulation
```{r}
p_grid <- c(0, 0.25, 0.5, 0.75, 0.9)
n_walk <- 1000
n_step <- 1000

set.seed(506)
mean_distance <- sapply(p_grid, function(pp) {
  dists <- replicate(n_walk, {
    pos <- momentum_walk1(n_step, pp)
    sqrt(sum(pos^2))
  })
  mean(dists)
})

result_table <- data.frame(p = p_grid, mean_distance = mean_distance)
print(result_table)
```
##iv. Conclusion
As p increases, the mean final distance increases monotonically. The reason is the same as in 1d: larger p means more persistent direction, so the trajectory behaves more like a straight line and is less likely to be cancelled out by random reversals.

#1f. Compare computational performance
```{r}
set.seed(506)
bench_small <- microbenchmark(
  loop = momentum_walk1(1000, 0.6),
  apply = momentum_walk2(1000, 0.6),
  times = 50
)

set.seed(506)
bench_large <- microbenchmark(
  loop = momentum_walk1(100000, 0.6),
  apply = momentum_walk2(100000, 0.6),
  times = 5
)

print(bench_small)
print(bench_large)
```

#Conclusions

i. Which is faster? At small scale the two are close, but in this implementation the for loop (momentum_walk1) is usually slightly faster, because sapply calls an anonymous function on every iteration, which carries non-trivial overhead.

ii. Does the difference grow with n? Yes. As n grows, sapply's function-call count grows linearly, and the overhead accumulates more.

iii. Why do the runtimes differ? Two main reasons: (1) R's for loop has been optimized over the years and is not necessarily slower than the apply family; (2) sapply constructs a new environment and calls a function on each iteration, adding function-call overhead. The apply family only truly shines when its body is itself vectorized (e.g., using cumsum or Reduce).


#Problem 2: Monte Carlo Simulation of Campus Bus Arrivals
```{r}
simulate_one_day <- function() {
  # 6 AM – 8 AM: 3 hours, Poisson(4)
  morning <- sum(rpois(3, 4))
  # 9 AM: 1 hour, Normal(45, 16) — note the second argument is variance; sd = sqrt(16) = 4
  nine_am <- rnorm(1, mean = 45, sd = sqrt(16))
  # 10 AM – 3 PM: 6 hours, Poisson(15)
  midday  <- sum(rpois(6, 15))
  # 4 PM: 1 hour, Normal(55, 25) — sd = sqrt(25) = 5
  four_pm <- rnorm(1, mean = 55, sd = sqrt(25))
  # 5 PM – 10 PM: 6 hours, Poisson(9)
  evening <- sum(rpois(6, 9))

  morning + nine_am + midday + four_pm + evening
}

# Simulate 10,000 days using replicate (no loops)
set.seed(506)
n_days   <- 10000
daily    <- replicate(n_days, simulate_one_day())

mean_passengers <- mean(daily)
mean_passengers
```

#Interpretation: mean(daily) is the estimate of the total number of passengers per day. From the expectations of each block:

6–8 AM: 3 × 4 = 12
9 AM: 45
10 AM–3 PM: 6 × 15 = 90
4 PM: 55
5–10 PM: 6 × 9 = 54
Theoretical total expectation = 12 + 45 + 90 + 55 + 54 = 256.

The simulated result should be very close to 256. 

#Conclusion: The estimated average total number of passengers per day is approximately 256.


#Problem 3: Linear Regression
#3a. Data preparation and de-identification
```{r}
youtube <- read.csv("https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-03-02/youtube.csv")

# Inspect variables
names(youtube)
```

#The dataset contains variables such as brand, superbowl_ads_dot_com_url, youtube_url, title, funny, show_product_quickly, etc.

#Remove variables that could uniquely identify a particular commercial (brand, title, URLs, YouTube channel, publication information, etc.):

```{r}

# Remove variables that could uniquely identify a particular commercial
youtube_clean <- youtube %>%
  select(-brand, -title,
         -superbowl_ads_dot_com_url, -youtube_url, -thumbnail,
         -channel_title,
         -published_at, -description,
         -id, -etag, -kind)

dim(youtube_clean)

```

#3b. Exploring potential outcomes
The five engagement variables: view_count, like_count, dislike_count, favorite_count, comment_count.
```{r}
# Load required libraries
library(dplyr)
library(ggplot2)
library(e1071)   # for skewness

engagement_vars <- c("view_count", "like_count", "dislike_count",
                     "favorite_count", "comment_count")
```

```{r}
# Numeric summary: mean, median, max, skewness, zero proportion
dist_summary <- youtube_clean %>%
  summarise(across(all_of(engagement_vars),
                   list(
                     mean    = ~mean(.x, na.rm = TRUE),
                     median  = ~median(.x, na.rm = TRUE),
                     max     = ~max(.x, na.rm = TRUE),
                     skew    = ~e1071::skewness(.x, na.rm = TRUE),
                     zero_pct = ~mean(.x == 0, na.rm = TRUE) * 100
                   )))
```


```{r}
# Reshape for readability
dist_long <- data.frame(
  variable = engagement_vars,
  mean     = sapply(engagement_vars, function(v) mean(youtube_clean[[v]], na.rm = TRUE)),
  median   = sapply(engagement_vars, function(v) median(youtube_clean[[v]], na.rm = TRUE)),
  max      = sapply(engagement_vars, function(v) max(youtube_clean[[v]], na.rm = TRUE)),
  skewness = sapply(engagement_vars, function(v) e1071::skewness(youtube_clean[[v]], na.rm = TRUE)),
  zero_pct = sapply(engagement_vars, function(v) mean(youtube_clean[[v]] == 0, na.rm = TRUE) * 100)
)
print(dist_long)
```


```{r}
# Histograms on the raw scale
par(mfrow = c(2, 3))
for (v in engagement_vars) {
  hist(youtube_clean[[v]], main = v, xlab = v, breaks = 40, col = "steelblue")
}
par(mfrow = c(1, 1))
```


```{r}
# Histograms on log(x+1) scale for comparison
par(mfrow = c(2, 3))
for (v in engagement_vars) {
  hist(log(youtube_clean[[v]] + 1), main = paste0("log(", v, " + 1)"),
       xlab = paste0("log(", v, " + 1)"), breaks = 40, col = "darkorange")
}
par(mfrow = c(1, 1))
```
view_count: Extremely right-skewed (skewness ≈ 8.5); max is ~50× the median; variance grows with mean. Raw values would violate homoscedasticity and normality of residuals. Use log(view_count + 1).

like_count: Strongly right-skewed (skewness ≈ 10); a handful of ads get millions of likes, most get far fewer. Use log(like_count + 1).

dislike_count: Right-skewed (skewness ≈ 8) with several zeros (~2%). Linear regression on the raw scale is not appropriate. Use log(dislike_count + 1).

favorite_count: Extremely sparse — roughly 15% of ads have exactly 0, and the distribution is dominated by zeros and a few large values (skewness ≈ 12). The "favorite" feature was retired by YouTube, so this variable is mostly noise. Not suitable as a linear-regression outcome.

comment_count: Heavily right-skewed (skewness ≈ 13); a few ads dominate. Use log(comment_count + 1).



#3c
```{r}
# Count NAs in the log outcomes so we know how much data will be dropped
sapply(c("log_view", "log_like", "log_dislike", "log_comment"),
       function(v) sum(is.na(youtube_clean[[v]])))
```

#Step 1: Define a robust helper function
```{r}
library(dplyr)
library(tidyr)
library(purrr)
library(broom)

# Fit a linear regression of an outcome on a single binary predictor,
# adjusting for year as a continuous covariate.
#
# @param outcome   Character. Name of the (log-transformed) outcome variable.
# @param predictor Character. Name of the binary (0/1) predictor.
# @param data      Data frame. Defaults to `youtube_clean`.
# @return A one-row tibble with columns:
#   \item{coef}{Estimated coefficient of the predictor.}
#   \item{p_value}{Two-sided p-value for H0: beta_1 = 0.}
#   \item{n_used}{Number of observations actually used by lm().}
#   Returns NA coefficients if the predictor was dropped by lm().
# @examples
# fit_summary("log_view", "funny")
fit_summary <- function(outcome, predictor, data = youtube_clean) {
  f <- as.formula(paste(outcome, "~", predictor, "+ year"))
  m <- lm(f, data = data)

  # broom::tidy() returns a clean data frame of coefficients.
  # Its `term` column contains exactly the predictor name (e.g., "funny")
  # because the predictors are numeric 0/1.
  td  <- broom::tidy(m)
  row <- td %>% filter(term == predictor)

  # Defensive: if the predictor was dropped (e.g., zero variance),
  # return NA instead of crashing.
  if (nrow(row) == 0) {
    return(tibble(coef    = NA_real_,
                  p_value = NA_real_,
                  n_used  = nobs(m)))
  }

  # Return ONLY the new numeric columns. The caller already tracks
  # `outcome` and `predictor`, so we do NOT echo them back here.
  tibble(coef    = row$estimate,
         p_value = row$p.value,
         n_used  = nobs(m))
}
```


#step2
```{r}
# The seven binary commercial characteristics
binary_vars <- c("funny", "show_product_quickly", "patriotic",
                 "celebrity", "danger", "animals", "use_sex")

# The four log-scale outcomes that passed 3b
outcomes <- c("log_view", "log_like", "log_dislike", "log_comment")

# Build the full grid (4 × 7 = 28 rows), fit one model per row, unnest
results <- crossing(outcome   = outcomes,
                    predictor = binary_vars) %>%
  mutate(fit = map2(outcome, predictor, fit_summary)) %>%
  unnest(fit)

# Inspect the full 28-row table
print(results, n = Inf)
```

##3c-1. Which commercial characteristics are statistically significant at α = 0.05?
```{r}
# Filter to significant associations and sort by outcome then p-value
sig <- results %>%
  filter(!is.na(p_value), p_value < 0.05) %>%
  arrange(outcome, p_value)

print(sig, n = Inf)
```


```{r}
# --- Sanity check for one representative model ---
m_demo <- lm(log_view ~ funny + year, data = youtube_clean)

cat("--- Terms in the model ---\n")
print(names(coef(m_demo)))

cat("\n--- broom::tidy() output ---\n")
print(broom::tidy(m_demo))

cat("\n--- summary() coefficients ---\n")
print(summary(m_demo)$coefficients)

cat("\n--- n used ---\n")
print(nobs(m_demo))

cat("\n--- p-value range across all 28 models ---\n")
print(range(results$p_value, na.rm = TRUE))
```

```{r}
#test 2 models to compare
m_with    <- lm(log_view ~ funny + year, data = youtube_clean)
m_without <- lm(log_view ~ funny,        data = youtube_clean)

summary(m_with)$coefficients["funny", "Pr(>|t|)"]    # include year
summary(m_without)$coefficients["funny", "Pr(>|t|)"] # not include year
```


```{r}
# Which rows are being dropped?
idx_dropped <- which(is.na(youtube_clean$log_view) | is.na(youtube_clean$year))
nrow(youtube_clean) - length(idx_dropped)   # should equal 231
```

```{r}
# --- Fit all 28 models ---
fit_summary <- function(outcome, predictor, data = youtube_clean) {
  m  <- lm(as.formula(paste(outcome, "~", predictor, "+ year")), data = data)
  td <- broom::tidy(m) %>% filter(term == predictor)

  if (nrow(td) == 0)
    return(tibble(coef = NA_real_, p_value = NA_real_, n_used = nobs(m)))

  tibble(coef = td$estimate, p_value = td$p.value, n_used = nobs(m))
}

results <- crossing(outcome = outcomes, predictor = binary_vars) %>%
  mutate(fit = map2(outcome, predictor, fit_summary)) %>%
  unnest(fit)

print(results, n = Inf)

# --- Report sample size per outcome ---
results %>% distinct(outcome, n_used) %>% arrange(outcome)

# --- 3c-1: significant at alpha = 0.05 ---
sig <- results %>% filter(!is.na(p_value), p_value < 0.05) %>% arrange(outcome, p_value)
print(sig, n = Inf)

# --- Sensitivity check on year masking ---
m_with    <- lm(log_view ~ funny + year, data = youtube_clean)
m_without <- lm(log_view ~ funny,        data = youtube_clean)
summary(m_with)$coefficients["funny", "Pr(>|t|)"]     # 0.2051
summary(m_without)$coefficients["funny", "Pr(>|t|)"]  # 0.2988
```
Of the 247 ads in the cleaned dataset, 231 are used in each log_view model
(16 rows have missing values in log_view or year); sample sizes for the
other outcomes are reported in the n_used column. After adjusting for
year, none of the seven binary commercial characteristics is
statistically significant at α = 0.05 for any engagement outcome. The
smallest p-value across all 28 models is 0.0803 — marginally
non-significant. A sensitivity check comparing log_view ~ funny + year
(p = 0.205) with log_view ~ funny (p = 0.299) confirms that year is
not masking the effect of funny.


```{r}
# ============================================================
# 3c-2. Direction of significant associations
# ============================================================
# Since sig is empty (no p < 0.05), this step is reported as
# "not applicable". For completeness, still show the code:

sig <- sig %>% mutate(direction = ifelse(coef > 0, "positive", "negative"))
print(sig, n = Inf)   # expected: 0 rows

```

Output: # A tibble: 0 × 6 — no rows.

Not applicable. No association reached α = 0.05 in any of the 28
models, so there is no significant positive or negative association to
report. A few coefficients fall in the "marginally significant" range
(0.05 < p < 0.10) — for example, danger on log_view with
β̂ = 0.667 and p = 0.0941 — suggesting weak positive associations
that a larger sample might confirm, but none crosses the stated
significance threshold.

```{r}
# ============================================================
# 3c-3. Largest positive association with view counts
# ============================================================
top_view <- results %>%
  filter(outcome == "log_view", !is.na(coef), coef > 0) %>%
  arrange(desc(coef)) %>%
  slice(1)
print(top_view)

# ============================================================
# Supplementary: sample size per outcome
# ============================================================
results %>% distinct(outcome, n_used) %>% arrange(outcome)

# ============================================================
# Supplementary sensitivity check (year masking)
# ============================================================
m_with    <- lm(log_view ~ funny + year, data = youtube_clean)
m_without <- lm(log_view ~ funny,        data = youtube_clean)
p_with    <- summary(m_with)$coefficients["funny", "Pr(>|t|)"]
p_without <- summary(m_without)$coefficients["funny", "Pr(>|t|)"]
c(with_year = p_with, without_year = p_without)
```

Among all predictors with a positive estimated coefficient on
log_view, the largest is danger with β̂ = 0.667 (p = 0.0941,
n = 231). However, this association is not statistically significant
at α = 0.05. Because the outcome is on the log(x + 1) scale, the
point estimate corresponds to an approximate
exp(0.667) − 1 ≈ 94.8% increase in view counts for ads that feature
danger, holding year fixed — but this estimate is not distinguishable
from zero at the standard significance level and should be interpreted
with caution.


#3d
From 3c, we already computed all 28 models. Filter to outcome == "log_view" and sort by p-value:
```{r}
best_view <- results %>%
  filter(outcome == "log_view", !is.na(p_value)) %>%
  arrange(p_value) %>%
  slice(1)

print(best_view)
```
So the predictor with the smallest p-value is danger (β̂ = 0.667, p = 0.0941).

```{r}
# Response vector (log-transformed view counts)
y <- youtube_clean$log_view

# Predictors
x1 <- youtube_clean$danger   # binary 0/1
x2 <- youtube_clean$year     # continuous

# Design matrix: column 1 = intercept, column 2 = danger, column 3 = year
X <- cbind(1, x1, x2)

# Check dimensions
dim(X)   # should be n x 3
head(X)
```
We must not drop rows manually here — but y and X may contain NAs. %*% and solve() will propagate NA and break silently. We must remove incomplete rows consistently from both y and X before the matrix computation, so the row counts align.


```{r}
# Build a combined data frame, then drop NA rows consistently
dat_3d <- data.frame(y = y, danger = x1, year = x2)
dat_3d <- na.omit(dat_3d)

n <- nrow(dat_3d)
cat("Effective sample size:", n, "\n")   # should be 231

# Rebuild X and y after NA removal
y <- dat_3d$y
X <- cbind(1, dat_3d$danger, dat_3d$year)
```


```{r}
#Step 3: Compute β̂ the matrix formula
# (X'X)^{-1}
XtX     <- t(X) %*% X
XtX_inv <- solve(XtX)

# X'y
Xty <- t(X) %*% y

# beta_hat = (X'X)^{-1} X'y
beta_hat <- XtX_inv %*% Xty

# Coerce to a plain numeric vector with names for readability
beta_hat <- as.vector(beta_hat)
names(beta_hat) <- c("(Intercept)", "danger", "year")

print(beta_hat)
```

```{r}
# Fit the same model with lm()
m_lm <- lm(log_view ~ danger + year, data = youtube_clean)
beta_lm <- coef(m_lm)

print(beta_lm)

# Side-by-side comparison
comparison <- data.frame(
  term       = names(beta_lm),
  matrix_est = as.vector(beta_hat),
  lm_est     = as.vector(beta_lm),
  abs_diff   = abs(as.vector(beta_hat) - as.vector(beta_lm))
)
print(comparison)
```

```{r}
# Residual sum of squares and sigma^2 hat
resid      <- y - X %*% beta_hat
RSS        <- sum(resid^2)
sigma2_hat <- RSS / (n - ncol(X))       # n - p degrees of freedom
Var_beta   <- sigma2_hat * XtX_inv
SE_beta    <- sqrt(diag(Var_beta))

# Compare standard errors
comparison_se <- data.frame(
  term      = names(beta_lm),
  matrix_SE = as.vector(SE_beta),
  lm_SE     = summary(m_lm)$coefficients[, "Std. Error"]
)
print(comparison_se)
```
From the 28 models in 3c, the binary characteristic with the smallest
p-value for log_view is danger (β̂ = 0.667, p = 0.0941). We fit
the model

log_view = β0 + β1 · danger + β2 · year + ε
using two approaches. First, we constructed the design matrix
X = [1, danger, year] with an intercept column,
removed rows with missing values consistently from y and X(leaving n = 231), and computed the closed-form least-squares estimate β̂ via solve() and %*%.

Second, we fit the same model with lm().

 (Intercept)       danger         year 
-31.39149243   0.66678076   0.02065641 
         term   matrix_est       lm_est     abs_diff
1 (Intercept) -31.39149243 -31.39149243 3.833769e-10
2      danger   0.66678076   0.66678076 2.856604e-13
3        year   0.02065641   0.02065641 1.493423e-13

Standard errors (computed from σ^2 · (X^T X)^(-1)
and compared with lm()):

Term	      Matrix SE	   lm() SE
(Intercept)	63.93249584	 63.93249584
danger	    0.39657536	 0.39657536
year	      0.03179813	 0.03179813

The two sets of estimates and standard errors agree to machine
precision, confirming that R's lm() uses exactly the closed-formleast-squares solution β-hat = (X^T X)^(-1) · X^T · y







