---
title: "homework2-506"
author: "Zhixuan Li"
format:
  html:
    embed-resources: true
editor: visual
---

Link to the repository: <https://github.com/lizhixuan202603-sudo/homework2-506>

```{r setup}
#| message: false
#| warning: false

library(microbenchmark)
library(dplyr)
library(tidyr)
library(purrr)
library(broom)
library(ggplot2)
library(e1071)

set.seed(506)
```

# Problem 1: Random Walk with Momentum

## 1a. Simulation function

### i. Function `momentum_walk(n, p)`
```{r}
#' Random Walk with Momentum in 2D
#'
#' Simulates a particle moving on a 2D grid. The first step is chosen
#' uniformly from four directions; each subsequent step preserves the
#' previous direction with probability `p`, and otherwise chooses a new
#' direction uniformly from the four directions.
#'
#' @param n Positive integer. Number of steps (n >= 1).
#' @param p Numeric. Probability of preserving the previous direction
#'   (0 <= p <= 1).
#' @return A numeric vector of length 2, c(x, y), the final coordinates.
#' @examples
#' momentum_walk(10, 0.6)
momentum_walk <- function(n, p) {
  if (!is.numeric(n) || length(n) != 1 || is.na(n)) stop("n must be a single numeric value")
  if (!is.numeric(p) || length(p) != 1 || is.na(p)) stop("p must be a single numeric value")
  if (n < 1 || n != as.integer(n)) stop("n must be an integer >= 1")
  if (p < 0 || p > 1) stop("p must be in [0, 1]")

  dirs <- matrix(c(1, 0,
                  -1, 0,
                   0, 1,
                   0,-1), ncol = 2, byrow = TRUE)

  d <- sample(1:4, 1)
  x <- dirs[d, 1]; y <- dirs[d, 2]
  prev <- d

  if (n > 1) {
    for (i in 2:n) {
      if (runif(1) < p) {
        d <- prev
      } else {
        d <- sample(1:4, 1)
      }
      x <- x + dirs[d, 1]
      y <- y + dirs[d, 2]
      prev <- d
    }
  }
  c(x, y)
}

momentum_walk(10, 0.6)
```

### ii. Input sanitization

The function validates `n >= 1` (and integer) and `0 <= p <= 1`, raising
an informative error otherwise.
```{r}
tryCatch(momentum_walk(0, 0.5),    error = function(e) message(e$message))
tryCatch(momentum_walk(10, 1.5),   error = function(e) message(e$message))
tryCatch(momentum_walk(10.5, 0.5), error = function(e) message(e$message))
```

## 1b. Two implementations

### Version 1: `for` loop
```{r}
#' Random Walk with Momentum — for-loop implementation
#'
#' @param n Number of steps.
#' @param p Probability of preserving direction.
#' @return Final coordinates c(x, y).
momentum_walk1 <- function(n, p) {
  if (!is.numeric(n) || length(n) != 1 || is.na(n)) stop("n must be a single numeric value")
  if (!is.numeric(p) || length(p) != 1 || is.na(p)) stop("p must be a single numeric value")
  if (n < 1 || n != as.integer(n)) stop("n must be an integer >= 1")
  if (p < 0 || p > 1) stop("p must be in [0, 1]")

  dirs <- matrix(c(1, 0, -1, 0, 0, 1, 0, -1), ncol = 2, byrow = TRUE)

  d    <- sample(1:4, 1)
  step <- dirs[d, ]
  x <- step[1]; y <- step[2]

  for (i in 2:n) {
    if (runif(1) < p) {
      d <- d
    } else {
      d <- sample(1:4, 1)
    }
    step <- dirs[d, ]
    x <- x + step[1]
    y <- y + step[2]
  }
  c(x, y)
}
```

**Logic.** The `for` loop advances the particle step-by-step. At each
step it first decides whether to keep the previous direction, then
updates the coordinates. The approach is intuitive, though loops in R
are relatively slow.

### Version 2: `sapply` implementation

```{r}
#' Random Walk with Momentum — sapply implementation
#'
#' @param n Number of steps.
#' @param p Probability of preserving direction.
#' @return Final coordinates c(x, y).
momentum_walk2 <- function(n, p) {
  if (!is.numeric(n) || length(n) != 1 || is.na(n)) stop("n must be a single numeric value")
  if (!is.numeric(p) || length(p) != 1 || is.na(p)) stop("p must be a single numeric value")
  if (n < 1 || n != as.integer(n)) stop("n must be an integer >= 1")
  if (p < 0 || p > 1) stop("p must be in [0, 1]")

  dirs <- matrix(c(1, 0, -1, 0, 0, 1, 0, -1), ncol = 2, byrow = TRUE)

  keep <- c(TRUE, runif(n - 1) < p)
  rand <- sample(1:4, n, replace = TRUE)

  idx <- integer(n)
  idx[1] <- rand[1]
  sapply(2:n, function(i) {
    if (keep[i]) {
      idx[i] <<- idx[i - 1]
    } else {
      idx[i] <<- rand[i]
    }
    idx[i]
  })

  xy <- dirs[idx, , drop = FALSE]
  c(sum(xy[, 1]), sum(xy[, 2]))
}
```

**Logic.** All "keep" decisions and candidate directions are generated
vectorially first; then `sapply` sequentially resolves each step's
direction; finally coordinates are summed in one shot, avoiding an
explicit `for` loop.

### Demonstration
```{r}
set.seed(1)
momentum_walk1(10, 0.6)
momentum_walk2(10, 0.6)
momentum_walk1(1000, 0.6)
momentum_walk2(1000, 0.6)
```


## 1c. Reproducibility
```{r}
set.seed(506)
r1 <- momentum_walk1(10, 0.6)
r2 <- momentum_walk1(10, 0.6)
set.seed(506)
r3 <- momentum_walk1(10, 0.6)

identical(r1, r3)
```

```{r}
set.seed(506)
m1 <- momentum_walk1(1000, 0.6)
set.seed(506)
m2 <- momentum_walk1(1000, 0.6)
identical(m1, m2)
```


**Explanation.** `set.seed()` sets the internal state of the
pseudo-random number generator. With the same seed, all subsequent random
calls produce an identical sequence, so results are reproducible.

## 1d. Does momentum matter?

### i. One simulation for each p

```{r}
set.seed(506)
p_vals <- c(0, 0.5, 0.9)
res_d  <- sapply(p_vals, function(pp) {
  pos <- momentum_walk1(1000, pp)
  sqrt(sum(pos^2))
})
names(res_d) <- paste0("p=", p_vals)
res_d
```

### ii. Larger momentum leads to larger final distance

The output above shows that D increases clearly as p increases.

### iii. Reasoning

When p = 0, each step is independent, so the particle performs a
classical 2D random walk; the variance of displacement grows linearly
with steps, and the final distance is roughly sqrt(n). When p approaches
1, the particle keeps moving in nearly the same direction, so
displacement grows nearly linearly (~n), and the final distance is much
larger than a random walk. Higher momentum means more "straight-line-like"
motion, so cancellations are less likely.

## 1e. Monte Carlo simulation

```{r}
p_grid <- c(0, 0.25, 0.5, 0.75, 0.9)
n_walk <- 1000
n_step <- 1000

set.seed(506)
mean_distance <- sapply(p_grid, function(pp) {
  dists <- replicate(n_walk, {
    pos <- momentum_walk1(n_step, pp)
    sqrt(sum(pos^2))
  })
  mean(dists)
})

data.frame(p = p_grid, mean_distance = mean_distance)
```

### iv. Conclusion

As p increases, the mean final distance increases monotonically. The
reason is the same as in 1d: larger p means more persistent direction,
so the trajectory behaves more like a straight line and is less likely
to be cancelled out by random reversals.

## 1f. Compare computational performance
```{r}
set.seed(506)
bench_small <- microbenchmark(
  loop  = momentum_walk1(1000, 0.6),
  apply = momentum_walk2(1000, 0.6),
  times = 50
)

set.seed(506)
bench_large <- microbenchmark(
  loop  = momentum_walk1(100000, 0.6),
  apply = momentum_walk2(100000, 0.6),
  times = 5
)

print(bench_small)
print(bench_large)
```

**Conclusions.**

i. **Which is faster?** In this implementation, the sapply version is faster than the for loop, and the gap widens as n grows. The reason is that the sapply version generates the random draws and the "keep" decisions vectorially first, and then resolves directions in a single apply pass, whereas the for loop pays R-level loop overhead on every iteration.

ii. **Does the difference grow with n?** Yes. As n grows, `sapply`'s
function-call count grows linearly, and the overhead accumulates more.

iii. **Why do the runtimes differ?** Two main reasons: (1) R's `for` loop
has been optimized over the years and is not necessarily slower than the
apply family; (2) `sapply` constructs a new environment and calls a
function on each iteration, adding function-call overhead. The apply
family only truly shines when its body is itself vectorized (e.g., using
`cumsum` or `Reduce`).


# Problem 2: Monte Carlo Simulation of Campus Bus Arrivals
```{r}
#' Simulate the total number of passengers boarding the bus in one day
#'
#' @return A single numeric value: the total number of passengers
#'   across all hours of one simulated day.
simulate_one_day <- function() {
  morning <- sum(rpois(3, 4))
  nine_am <- rnorm(1, mean = 45, sd = sqrt(16))
  midday  <- sum(rpois(6, 15))
  four_pm <- rnorm(1, mean = 55, sd = sqrt(25))
  evening <- sum(rpois(6, 9))
  morning + nine_am + midday + four_pm + evening
}

set.seed(506)
n_days <- 10000
daily  <- replicate(n_days, simulate_one_day())

mean(daily)
```

**Interpretation.** From the expectations of each block:

- 6-8 AM: 3 x 4 = 12
- 9 AM: 45
- 10 AM-3 PM: 6 x 15 = 90
- 4 PM: 55
- 5-10 PM: 6 x 9 = 54

Theoretical total expectation = 12 + 45 + 90 + 55 + 54 = **256**.

**Conclusion.** The estimated average total number of passengers per day
is approximately **256**, matching the theoretical expectation.


# Problem 3: Linear Regression

## 3a. Data preparation and de-identification
```{r}
youtube <- read.csv(
  "https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2021/2021-03-02/youtube.csv"
)
```

We remove variables that uniquely identify each commercial
(`brand`, `title`, URLs, `channel_title`, publication info, etc.):

```{r}
youtube_clean <- youtube %>%
  select(-brand, -title,
         -superbowl_ads_dot_com_url, -youtube_url, -thumbnail,
         -channel_title,
         -published_at, -description,
         -id, -etag, -kind)

dim(youtube_clean)
```

**Result.** The cleaned dataset has **247 rows** and **14 columns**.

## 3b. Exploring potential outcomes

The five engagement variables are:

- `view_count`
- `like_count`
- `dislike_count`
- `favorite_count`
- `comment_count`

```{r}
engagement_vars <- c("view_count", "like_count", "dislike_count",
                     "favorite_count", "comment_count")

dist_long <- data.frame(
  variable = engagement_vars,
  mean     = sapply(engagement_vars, function(v) mean(youtube_clean[[v]],   na.rm = TRUE)),
  median   = sapply(engagement_vars, function(v) median(youtube_clean[[v]], na.rm = TRUE)),
  max      = sapply(engagement_vars, function(v) max(youtube_clean[[v]],    na.rm = TRUE)),
  skewness = sapply(engagement_vars, function(v) e1071::skewness(youtube_clean[[v]], na.rm = TRUE)),
  zero_pct = sapply(engagement_vars, function(v) mean(youtube_clean[[v]] == 0, na.rm = TRUE) * 100)
)
dist_long
```

```{r}
par(mfrow = c(2, 3))
for (v in engagement_vars) {
  hist(youtube_clean[[v]], main = v, xlab = v, breaks = 40, col = "steelblue")
}
par(mfrow = c(1, 1))

par(mfrow = c(2, 3))
for (v in engagement_vars) {
  hist(log(youtube_clean[[v]] + 1), main = paste0("log(", v, " + 1)"),
       xlab = paste0("log(", v, " + 1)"), breaks = 40, col = "darkorange")
}
par(mfrow = c(1, 1))
```


### Classification of each outcome

view_count — (2) Transform. Extremely right-skewed (skewness = 13.63, zero-percent = 0%); the maximum (1.76e8) is roughly 4000x the median (4.14e4). Use log(view_count + 1).

like_count — (2) Transform. Strongly right-skewed (skewness = 8.66, zero-percent = 4.00%); a handful of ads get hundreds of thousands of likes while most get far fewer. Use log(like_count + 1).

dislike_count — (2) Transform. Right-skewed (skewness = 11.40, zero-percent = 20.89%). Use log(dislike_count + 1).

favorite_count — (3) Not appropriate. Every value is exactly 0 (zero-percent = 100%), so the skewness is undefined (NaN). Because the variable has no variation at all, it cannot be used as a regression outcome. (The "favorite" feature was retired by YouTube, so this column carries no information.)

comment_count — (2) Transform. Heavily right-skewed (skewness = 7.88, zero-percent = 18.47%). Use log(comment_count + 1).
  
### Apply all data prep: binary coercion + log transform

This is the single chunk that prepares everything for 3c and 3d.

```{r}
binary_vars <- c("funny", "show_product_quickly", "patriotic",
                 "celebrity", "danger", "animals", "use_sex")

youtube_clean <- youtube_clean %>%
  mutate(across(all_of(binary_vars), function(x) as.numeric(as.logical(x)))) %>%
  mutate(
    log_view    = log(view_count + 1),
    log_like    = log(like_count + 1),
    log_dislike = log(dislike_count + 1),
    log_comment = log(comment_count + 1)
  )

str(youtube_clean[c(binary_vars,
                    "log_view", "log_like", "log_dislike", "log_comment")])

sapply(c("log_view", "log_like", "log_dislike", "log_comment"),
       function(v) e1071::skewness(youtube_clean[[v]], na.rm = TRUE))

stopifnot(all(c("log_view", "log_like", "log_dislike", "log_comment")
              %in% names(youtube_clean)))
stopifnot(all(binary_vars %in% names(youtube_clean)))
```

**Interpretation.** log_view, log_like, and log_comment are now approximately symmetric (|skewness| < 0.7). log_dislike retains mild right-skew (skewness = 1.33) but is dramatically improved over the raw scale. All seven binary predictors are now numeric 0/1. favorite_count is dropped entirely: it is constant at 0, so it carries no information and cannot serve as a regression outcome.

## 3c. Fitting linear regression models

### Setup: outcomes, helper function

```{r}
outcomes <- c("log_view", "log_like", "log_dislike", "log_comment")

#' Fit a linear regression of an outcome on one binary predictor,
#' adjusting for year as a continuous covariate.
#'
#' @param outcome   Character. Name of the (log-transformed) outcome.
#' @param predictor Character. Name of the binary (0/1) predictor.
#' @param data      Data frame. Defaults to `youtube_clean`.
#' @return A one-row tibble with columns `coef`, `p_value`, `n_used`.
#' @examples
#' fit_summary("log_view", "funny")
fit_summary <- function(outcome, predictor, data = youtube_clean) {
  m  <- lm(as.formula(paste(outcome, "~", predictor, "+ year")), data = data)
  td <- broom::tidy(m) %>% filter(term == predictor)

  if (nrow(td) == 0)
    return(tibble(coef = NA_real_, p_value = NA_real_, n_used = nobs(m)))

  tibble(coef    = td$estimate,
         p_value = td$p.value,
         n_used  = nobs(m))
}
```

### Fit all 28 models
```{r}
results <- crossing(outcome = outcomes, predictor = binary_vars) %>%
  mutate(fit = map2(outcome, predictor, fit_summary)) %>%
  unnest(fit)

results
```


### Sample size per outcome
```{r}
results %>% distinct(outcome, n_used) %>% arrange(outcome)
```

### 3c-1. Significant characteristics at alpha = 0.05
```{r}
sig <- results %>%
  filter(!is.na(p_value), p_value < 0.05) %>%
  arrange(outcome, p_value)
sig
```

**Written answer.** Of the 247 ads in the cleaned dataset, 231 are used in every log_view model (16 rows have missing values in log_view or year); the other outcomes use 225 (log_like, log_dislike) and 222 (log_comment) observations. favorite_count is excluded entirely because it is constant at 0.

After adjusting for year, none of the seven binary commercial characteristics is statistically significant at alpha = 0.05 for any of the four engagement outcomes. The smallest p-value across all 28 models is 0.0941 (for log_view ~ danger), which is marginally non-significant.

**Sensitivity check.** To verify that `year` is not masking the effect of
`funny`, we compare `lm(log_view ~ funny + year)` with `lm(log_view ~
funny)`:

```{r}
m_with    <- lm(log_view ~ funny + year, data = youtube_clean)
m_without <- lm(log_view ~ funny,        data = youtube_clean)

c(with_year    = summary(m_with)$coefficients["funny", "Pr(>|t|)"],
  without_year = summary(m_without)$coefficients["funny", "Pr(>|t|)"])
```

The p-value for `funny` is **0.205** with year and **0.299** without
year. Removing `year` does not reveal a hidden association. This
confirms that the non-significance reflects genuine weak signal in the
data rather than a modeling artifact.

### 3c-2. Direction of significant associations
```{r}
sig <- sig %>% mutate(direction = ifelse(coef > 0, "positive", "negative"))
sig
```

**Written answer.** **Not applicable.** No association reached alpha = 0.05 in any of the 28 models, so there is no significant positive or negative association to report. A few coefficients fall in the "marginally significant" range (0.05 < p < 0.10) — for example, danger on log_view with beta-hat = 0.667 and p = 0.0941 — suggesting a weak positive association that a larger sample might confirm, but none crosses the stated significance threshold.

### 3c-3. Largest positive association with view counts

```{r}
top_view <- results %>%
  filter(outcome == "log_view", !is.na(coef), coef > 0) %>%
  arrange(desc(coef)) %>%
  slice(1)
top_view
```

**Written answer.** Among all predictors with a positive estimated coefficient on log_view, the largest is danger with beta-hat = 0.667 (p = 0.0941, n = 231). However, this association is not statistically significant at alpha = 0.05. Because the outcome is on the log(x + 1) scale, the point estimate corresponds to an approximate exp(0.667) - 1 = 94.8% increase in view counts for ads that feature danger, holding year fixed. But this estimate is not distinguishable from zero at the standard significance level, so it should be reported as suggestive rather than conclusive.

## 3d. Estimating regression coefficients using matrix algebra

### Select the predictor with the smallest p-value for `log_view`

```{r}
best_view <- results %>%
  filter(outcome == "log_view", !is.na(p_value)) %>%
  arrange(p_value) %>%
  slice(1)
best_view
```

The predictor with the smallest p-value is **`danger`** (beta-hat = 0.667,
p = 0.0941).

### Build y and X, drop NAs consistently

```{r}
dat_3d <- data.frame(
  y      = youtube_clean$log_view,
  danger = youtube_clean$danger,
  year   = youtube_clean$year
)
dat_3d <- na.omit(dat_3d)

n_3d <- nrow(dat_3d)
cat("Effective sample size:", n_3d, "\n")

y_3d <- dat_3d$y
X_3d <- cbind(1, dat_3d$danger, dat_3d$year)
```

### Compute beta-hat via the matrix formula
```{r}
XtX_inv  <- solve(t(X_3d) %*% X_3d)
Xty      <- t(X_3d) %*% y_3d
beta_hat <- as.vector(XtX_inv %*% Xty)
names(beta_hat) <- c("(Intercept)", "danger", "year")
beta_hat
```

### Fit the same model with `lm()` and compare

```{r}
m_lm    <- lm(log_view ~ danger + year, data = youtube_clean)
beta_lm <- coef(m_lm)

comparison <- data.frame(
  term       = names(beta_lm),
  matrix_est = beta_hat,
  lm_est     = as.vector(beta_lm),
  abs_diff   = abs(beta_hat - as.vector(beta_lm))
)
comparison
```

### Compare standard errors
```{r}
resid      <- y_3d - X_3d %*% beta_hat
RSS        <- sum(resid^2)
sigma2_hat <- RSS / (n_3d - ncol(X_3d))
SE_beta    <- sqrt(diag(sigma2_hat * XtX_inv))

comparison_se <- data.frame(
  term      = names(beta_lm),
  matrix_SE = as.vector(SE_beta),
  lm_SE     = summary(m_lm)$coefficients[, "Std. Error"]
)
comparison_se
```

```{r}
all.equal(as.vector(beta_hat), as.vector(coef(m_lm)), tolerance = 1e-8)
```


### Written answer

From the 28 models in 3c, the binary characteristic with the smallest p-value for log_view is danger (beta-hat = 0.667, p = 0.0941). We fit the model

log_view = beta_0 + beta_1 * danger + beta_2 * year + epsilon

using two approaches. First, we constructed the design matrix X = [1, danger, year] with an intercept column, removed rows with missing values consistently from y and X (leaving n = 231), and computed the closed-form least-squares estimate beta-hat = (X'X)^(-1) X'y via solve() and %*%. Second, we fit the same model with lm().

Estimated coefficients:

Term	       Matrix estimate	 lm() estimate	  Absolute difference
(Intercept)	-31.39149243	     -31.39149243	    3.83e-10
danger	     0.66678076	       0.66678076	    2.86e-13
year	       0.02065641	       0.02065641	    1.49e-13

Standard errors computed from sigma^2 * (X'X)^(-1) also match lm()'s standard errors to machine precision:

Term	       Matrix SE	      lm() SE
(Intercept)	 63.93249584	    63.93249584
danger	     0.39657536	      0.39657536
year	       0.03179813	      0.03179813

The two sets of estimates and standard errors agree to machine precision, confirming that R's lm() uses exactly the closed-form least-squares solution beta-hat = (X'X)^(-1) X'y.




